Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-12 chapters/paper3.tex
Changes 1. Modified conclusion section in final remarks 2. Changes image sized in structural response of bijels 3. modified abstract and introduction to include structural analysis results
Changes 1. Modified rheology of bijels chapter

\begin{figure} 
    \centering 
    \includegraphics[scale=0.5]{../figures/results/paper3/stress-time_compare.png} 
    \caption{Time evolution of the shear stress of bijels stabilized with ellipsoidal particles as a function of the applied strain rate as
             a function of the initial microstructure and applied magnetic field. The observed stress response is difficult to say when a steady
             state is reached.} 
    \label{fig:stress_time} 
\end{figure}

@@

Bonaccorso et al. identified that the spherical particle stabilizers in their bijel simulations oriented to the direction of shear before being
ejected from the interface. \cite{bonaccorso_shear_2020} Due to the shape of ellipsoidal particles, they adopt a preferred orientation which
maximizes the interface covered by the particle. \cite{davies_dipolar_2015, davies_interface_2014} Energy must be imparted to rotate the particles
out of the interface. Upon application of shear, the orientation of particles on the interface will be a function of the applied shear rate, 
applied magnetic field and the capillary forces of the interface on the particle. \cite{cao_modeling_2021, chhabra_drag_1999, brenner_stokes_1963, naga_capillary_2021}
Therefore, the degree of interface angle change observed can shed light onto the rheological 
differences of bijels stabilized by prolate and oblate particles as the shear stress determined is affected by the particle rotation on the interface.

We analyzed the angle between each particles symmetry axis and the local interface normal by generating a mesh of the interface using the marching 
cubes algorithm and, for each particle, identified the nearest mesh vertex (excluding those beyond the length of the particle's long axis). We then calculated 
the angle $\psi$ between the particle orientation vector \(\hat{\vec{o}}_i\) and the local surface normal. The time evolution of the average interface
angle of all particles $\langle \psi \rangle$, calculated as a numerical average of $\psi$ for all particles, is shown in Figure~\ref{fig:interface_angle_shear}.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/psi-time_compare.png} 
    \caption{Comparisons of the interfacial angle of bijels stabilized by ellipsoidal particles as a function of 
             the applied shear rate at four different processing histories. We show that upon application of shear
             particles tilt out of the interface reduces the interfacial area stabilized. $\bar{B}: 0 \to 1$ shows 
             large variations in the interface angle.} 
    \label{fig:interface_angle_shear} 
\end{figure}

As shown in Figure~\ref{fig:interface_angle_shear}, the interface angle $\langle \psi \rangle$ generally increases over time across all cases, with the 
exception of the $\bar{B}: 0 \rightarrow 1$ condition. This indicates shear induced particle reorientation at the interface, indicating that particles are
being pulled out of the interface. The exception at $\bar{B}: 0 \rightarrow 1$ is consistent with the findings presented in Chapter~\ref{chapter:aim2}, where the 
observed behavior was attributed to particle reorientation toward the applied magnetic field. In all other cases, the gradual increase in $\langle \psi \rangle$ 
suggests that particles are progressively reorienting out of the interface under the influence of shear. For oblate particles initialized from bijel templates 
generated under $\bar{B} = 1$, the average initial interface angle is approximately 40°, indicating a pronounced tilt at the interface. As the tilt angle 
increases, the energy barrier to rotate a particle further out of the interface decreases, leading to reduced hydrodynamic drag and, consequently, lower 
effective viscosity. In contrast, the difference in initial interface angle between field- and non-field-aligned bijels stabilized by prolate particles 
is much smaller—approximately 13.5° and 11.5°, respectively. This smaller variation helps explain the more modest changes in rheological response observed 
in prolate systems, compared to the more pronounced field-dependent behavior in oblate particle-stabilized bijels.

@@

In this study, we investigated the rheological behavior of bijels stabilized by magnetically responsive ellipsoidal particles subjected to constant shear. 
In systems like solvent transfer-induced phase separation (STrIPS) and volatile-induced phase separation (VIPS), rheology plays a key role in determining the
microstructure and morphology of the bijel. The presence of anisotropic particles and external magnetic fields further complicates this behavior 
by altering interfacial structure. To systematically probe these effects, we employed a hybrid Lattice Boltzmann-Molecular Dynamics (LBM-MD) simulation 
framework to evaluate how particle shape, magnetic field history, and interfacial microstructure jointly influence the flow response of bijels. 
Four magnetic processing pathways were considered—ranging from fully unordered systems to those with persistent particle alignment—across a range of 
shear capillary numbers. By fitting the resulting stress-strain data to a Herschel-Bulkley model, we characterized the yield stress and degree of shear 
thinning for each configuration.

@@

To interpret these rheological differences, we analyzed both local structural ordering (via the Steinhardt $Q_6$ parameter) and particle orientation relative to 
the interface. Increased magnetic alignment in prolate systems led to higher $Q_6$ values and lower interface angles, indicating that particles remained 
well-anchored to the interface. This ordering enhanced stress distribution, leading to higher yield stress. In contrast, oblate particles 
tended to tilt further from the interface when aligned, as confirmed by elevated interface angles and decreased $Q_6$, making the monolayer more susceptible to 
shear-induced buckling and collapse. The increased interface angle of oblate particles after application of a field reduces the torque necessary to rotate particles
out of the interface, resulting in the larger differences in shear stress calculated for bijels stabilized with oblate particles.

These findings highlight the interplay between particle anisotropy, interfacial order, and external magnetic fields in tuning the mechanical properties of bijels. 
By modulating particle alignment through magnetic stimuli, it is possible to tailor the yield stress and flow behavior of these materials—a property of potential 
interest in fields such as soft robotics, drug delivery, and responsive templating. Moreover, the contrast in behavior between oblate and prolate particles 
underscores the importance of shape-specific design in engineering structured fluids. In the next chapter, we extend this investigation to consider dynamic or 
transient shear regimes and explore the reversibility of these structural transitions under cyclic loading and field modulation.

>>>>

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/stress-time_compare.png} 
    \caption{Time evolution of the shear stress of bijels stabilized with ellipsoidal particles as a function of the applied strain rate as
             a function of the initial microstructure and applied magnetic field. The observed stress response is difficult to say when a steady
             state is reached.} 
    \label{fig:stress_time} 
\end{figure}

@@

Bonaccorso et al. identified that the spherical particle stabilizers in their bijel simulations oriented to the direction of shear before being
ejected from the interface. \cite{bonaccorso_shear_2020} Due to the shape of ellipsoidal particles, they adopt a preferred orientation which
maximizes the interface covered by the particle. \cite{davies_dipolar_2015, davies_interface_2014} Energy must be imparted to rotate the particles
out of the interface. Upon application of shear, the orientation of particles on the interface will be a function of the applied shear rate, 
applied magnetic field and the capillary forces of the interface on the particle. \cite{cao_modeling_2021, chhabra_drag_1999, brenner_stokes_1963, naga_capillary_2021}
Therefore, the degree of interface angle change observed can shed light onto the rheological 
differences of bijels stabilized by prolate and oblate particles as the shear stress determined is affected by the particle rotation on the interface. We focus on
the effect of particle alignment to the shear direction to isolate the effect of shear driven particle realignment at the interface.

To assess the particle alignment to the applied shear, we calculate the average angle to the direction of shear, $\langle \theta_z \rangle$. This is calculated
by calculating the angle between the orientation vector $\vec{o}$ and the z axis $\vec{r_z}$ given by 
$\theta_z = \arccos(\frac{ \vec{o} \cdot \vec{r_z} }{ |\vec{o}| |\vec{r_z}| })$. This is done for all particles with $\langle \theta_z \rangle$ calculated from
by taking the numerical average of $\theta_z$ for all particles. The results are plotted in Figure \ref{fig:theta_z_specific}.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/theta_z-time_compare-specific.png} 
    \caption{Comparisons of the interfacial angle of bijels stabilized by ellipsoidal particles as a function of 
             the applied shear rate at four different processing histories. We show that upon application of shear
             particles tilt out of the interface reduces the interfacial area stabilized. $\bar{B}: 0 \to 1$ shows 
             large variations in the interface angle.} 
    \label{fig:theta_z_specific} 
\end{figure}

From Figure \ref{fig:theta_z_specific}, the effect of the aspect ratio of the particle stabilizer and the ordering of particles to the magnetic field. For oblate 
particles, the change in $\langle \theta_z \rangle$ over time can be seen to stay approximately the same at shear rates below $Ca = 1e-4$ However at higher shear rates,
$\langle \theta_z \rangle$ increases rapidly. When comparing this to the results when beginning with magnetic field ordering, $\langle \theta_z \rangle$ can be seen to
stay approximately the same only at $Ca = 1e-6$. This indicates shear driven reorientation of particles at the interface occurs at higher shear rates in bijels stabilized
with disordered particles and matching the trend observed in the yield stress. The results for the prolate particles show that particle order increases the applied shear
necessary to cause particle reorientation on the interface. When taking the $\langle Q6 \rangle$ results into context from Figure \ref{fig:q6_rheology_init}, 
the effect of magnetic field driven ordering of particles can be seen.

As the particles are adsorbed onto the interface, there is an open question whether the reorientation of particles to the shear also causes the interface to move with the
particles. To analyze this effect, we analyzed the angle between each particles symmetry axis and the local interface normal by generating a mesh of the interface using the marching 
cubes algorithm and, for each particle, identified the nearest mesh vertex (excluding those beyond the length of the particle's long axis). We then calculated 
the angle $\psi$ between the particle orientation vector \(\hat{\vec{o}}_i\) and the local surface normal. The time evolution of the average interface
angle of all particles $\langle \psi \rangle$, calculated as a numerical average of $\psi$ for all particles, is shown in Figure~\ref{fig:interface_angle_shear}.

\begin{figure} 
    \centering 
    \includegraphics[scale=0.3]{../figures/results/paper3/psi-time_compare-specific.png} 
    \caption{Comparisons of the interfacial angle of bijels stabilized by ellipsoidal particles as a function of 
             the applied shear rate at four different processing histories. We show that upon application of shear
             particles tilt out of the interface reduces the interfacial area stabilized.} 
    \label{fig:interface_angle_shear} 
\end{figure}

The interfacial angle results reveal how particle orientation relative to the local interface evolves under varying shear conditions, highlighting the interplay between particle shape, 
initial ordering, and shear rate. For disc-like particles, $\langle \psi \rangle$ increases steadily with time, especially at higher shear rates, indicating that shear 
flow drives discs to tilt further away from the interface normal. This effect is strengthened when the particles are initially ordered as the particles are easier to rotate out of the
interface when they are tilted. In contrast, rod-like particles exhibit more modest changes in $\langle \psi \rangle$ as the particles are closer to lying flat on the interface, meaning
that higher shear rates are needed to tilt them out of the interface. This effect can be seen more clearly when comparing the results for ordered disc and rod like particles, with
rod like particles demonstrating smaller average interfacial tilting than disc like particles.

Comparing the trends in $\langle \psi \rangle$ and $\langle \theta_z \rangle$ reveals a complementary view of how shear flow governs particle reorientation in bijels. For systems 
stabilized with oblate particles under no magnetic field and prolate particles subjected to a magnetic field, the results show that at low shear rates, particles largely resist 
reorientation due to interfacial jamming and anchoring forces. However, as the shear rate increases, both $\langle \psi \rangle$ and $\langle \theta_z \rangle$ rise, indicating that 
particles begin to tilt away from both the interface normal and the shear direction. This suggests that while the particles remain adsorbed at the interface, the imposed shear is strong 
enough to overcome local resistance and drive reorientation. The resulting changes in local packing and alignment are reflected in variations in the bond-orientational order parameter 
$\langle Q6 \rangle$, particularly under conditions where the bijel retains structural integrity and resists complete flow-induced breakdown.

@@

In this study, we investigated the rheological behavior of bijels stabilized by magnetically responsive ellipsoidal particles subjected to constant shear. 
In systems like Solvent Transfer Induced Phase Separation (STrIPS) and Vapor-Induced Phase Separation (VIPS), rheology plays a key role in determining the
microstructure and morphology of the bijel. The presence of anisotropic particles and external magnetic fields further complicates this behavior 
by altering interfacial structure. To systematically probe these effects, we employed a hybrid Lattice Boltzmann-Molecular Dynamics (LBM-MD) simulation 
framework to evaluate how particle shape, magnetic field history, and interfacial microstructure jointly influence the flow response of bijels. 
Four magnetic processing pathways were considered—ranging from fully unordered systems to those with persistent particle alignment—across a range of 
shear capillary numbers. By fitting the resulting stress-strain data to a Herschel-Bulkley model, we characterized the yield stress and degree of shear 
thinning for each configuration.

@@

To interpret these rheological differences, we analyzed the local structural ordering via the Steinhardt $Q_6$ parameter, and showed that increased magnetic alignment in 
prolate systems led to higher $Q_6$ values which corresponds to greater friction between particles, which manifests as a greater yield stress. The opposite trend was
characterized for oblate particles which explains the yield stress decrease as the particle order increases. To understand the effect of particle ordering 
on the yielding behavior of the bijel better, we focused on the $\bar{B}:0 \rightarrow 0$ $\bar{B}:1 \rightarrow 1$. We showed from the angle of the particles to the applied
shear, that bijels with greater local ordering resist shear better, as characterized by the lack of changes seen at lower shear rates compared to those with lower local
particle order. At higher shear rates, the angle to the applied shear changes. The interface angle provides a complementary view of the interfacial dynamics of the particle 
under shear, providing further evidence that the orientation of particles under shear corresponds to the local particle ordering. We also show that when the particles begin
to reorient to the applied shear, the interface moves as well.

These findings highlight the interplay between particle anisotropy, interfacial order, and external magnetic fields in tuning the mechanical properties of bijels. 
By modulating particle alignment through magnetic stimuli, it is possible to tailor the yield stress and flow behavior of these materials relevant in fields such as soft 
robotics, drug delivery, and responsive templating. Moreover, the contrast in behavior between oblate and prolate particles 
underscores the importance of particles in engineering the structure of bijels.
